---
title:"Temporal biotic homogenisation patterns vary with taxa"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
  rmdformats::html_clean:
    highlight: kate
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(cache = TRUE)
```


# Data management

## Package
Load packages
```{r}
pacman::p_load(
  tidyverse,                   # data management
  lme4, lmerTest, broom,       # GLM
  ggeffects, patchwork, ggsci, gridExtra, # ggplot
  vegan, iNEXT, betapart       # composition
  )
```


## load data
```{r message=FALSE, warning=FALSE}
# plant
df_plant <- read_csv(
  "data/df_plant_output.csv", 
  locale = locale(encoding = "cp932"),
  show_col_types = FALSE
  )

# bird
df_bird <- read_csv(
  "data/df_bird_output.csv", 
  locale = locale(encoding = "cp932"),
  show_col_types = FALSE
  )

# butterfly
df_butterfly <- read_csv(
  "data/df_butterfly_output.csv", 
  locale = locale(encoding = "cp932"),
  show_col_types = FALSE
  )

# environment
df_env <- read_csv(
  "data/df_env_output.csv", 
  locale = locale(encoding = "cp932"),
  show_col_types = FALSE
  )
```

# Fig. 1 
Comparison of regional species richness between past and present using rare faction curves

## Rarafaction curve
Function of iNEXT 
```{r}
est_inext <- function(x) {
  
  past <- x %>% 
    dplyr::select(site, time, sciname, value) %>% 
    dplyr::filter(time == "past") %>% 
    pivot_wider(names_from = site, values_from = value) %>% 
    dplyr::mutate(across(everything(), ~ ifelse(is.na(.), 0, .))) %>% 
    dplyr::filter(rowSums(across(where(is.numeric)), na.rm = TRUE) >= 0) %>% 
    column_to_rownames("sciname") %>% 
    dplyr::select(-time) 
  
  present <- x %>% 
    dplyr::select(site, time, sciname, value) %>% 
    dplyr::filter(time == "now") %>% 
    pivot_wider(names_from = site, values_from = value) %>% 
    dplyr::mutate(across(everything(), ~ ifelse(is.na(.), 0, .))) %>% 
    dplyr::filter(rowSums(across(where(is.numeric)), na.rm = TRUE) > 0) %>% 
    column_to_rownames("sciname") %>% 
    dplyr::select(-time) 
  
  
  # list for inext
  list_inext <- list("Past" = past, "Present" = present) 
  
  
  # doing 
  iNEXT(list_inext, q = 0, datatype = "incidence_raw")
  
}
```

## Run iNext
```{r}
## Plant
df_plant %>% {
  
  # All species
  est_inext(.) ->> inext_plant
  
  # Native species
  dplyr::filter(., exotic == "Native") %>% 
    est_inext(.) ->> inext_plant_n
  
  }


## Bird
df_bird %>% {
  
  # All species
  est_inext(.) ->> inext_bird
  
  # Native species
  dplyr::filter(., exotic == "Native") %>% 
    est_inext(.) ->> inext_bird_n
  }


## Butterfly
df_butterfly %>% {
  
  # All species
  est_inext(.) ->> inext_butterfly
  
  # Native species
  dplyr::filter(., exotic == "Native") %>% 
    est_inext(.) ->> inext_butterfly_n
  }
```

## Plot rare factrion curve
```{r message=FALSE, warning=FALSE}
# Plant
p_inext_plant <-
  ggiNEXT(inext_plant) +
    labs(
      title = "Plant",
      y = "All species richness"
      ) +
    scale_color_manual(values = c("#00AFBB", "#E7B800")) +
    scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
    scale_x_continuous(limits = c(0, 30)) +
    theme_bw(base_size = 8) +
    theme(
      panel.grid = element_blank(),
      axis.title.x = element_blank(),
      plot.subtitle = element_text(hjust = 0.5, size = 10),
      legend.position = "none"
      )


# Native plant
p_inext_plant_n <-
  ggiNEXT(inext_plant_n) +
    labs(
      x = "",
      y = "Native species richness"
      ) +
    scale_color_manual(values = c("#00AFBB", "#E7B800")) +
    scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
    scale_x_continuous(limits = c(0, 30)) +
    theme_bw(base_size = 8) +
    theme(
      panel.grid = element_blank(),
      plot.subtitle = element_text(hjust = 0.5, size = 10),
      legend.position = "none"
      )




# Bird
p_inext_bird <-
  ggiNEXT(inext_bird) +
    labs(title = "Bird") +
    scale_color_manual(values = c("#00AFBB", "#E7B800")) +
    scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
    scale_x_continuous(limits = c(0, 15)) +
    theme_bw(base_size = 8) +
    theme(
      panel.grid = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      legend.position = "none"
      )



# Native bird
p_inext_bird_n <-
  ggiNEXT(inext_bird_n) +
    labs(x = "Number of sites") +
    scale_color_manual(values = c("#00AFBB", "#E7B800")) +
    scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
    scale_x_continuous(limits = c(0, 15)) +
    theme_bw(base_size = 8) +
    theme(
      panel.grid = element_blank(),
      axis.title.y = element_blank(),
      legend.position = "none"
      )



# Butterfly
p_inext_butterfly <-
  ggiNEXT(inext_butterfly) +
    labs(title = "Butterfly") +
    scale_color_manual(values = c("#00AFBB", "#E7B800")) +
    scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
    scale_x_continuous(limits = c(0, 15)) +
    theme_bw(base_size = 8) +
    theme(
      panel.grid = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      legend.position = "none"
      )


# Native butterfly
p_inext_butterfly_n <-
  ggiNEXT(inext_butterfly_n) +
  labs(
    x = "",
    y = "Species richness"
  ) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
  scale_x_continuous(limits = c(0, 15)) +
  theme_bw(base_size = 8) +
  theme(
    panel.grid = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
    )
```

Plot combine
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
# Extract a legend
p_legend <- inext_butterfly_n %>% 
  ggiNEXT() + 
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
  theme(
    legend.direction = "vertical",
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.5, 'cm'))

legend_inext <- p_legend %>% 
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot()


# Axis label
grid.arrange(
  p_inext_plant,  p_inext_bird, p_inext_butterfly,
  p_inext_plant_n, p_inext_bird_n, p_inext_butterfly_n,
  legend_inext,
  nrow = 2,
  layout_matrix = rbind(c(1, 2, 3, 7),
                        c(4, 5, 6, 7)),
  widths = c(1.05, 1, 0.95, 0.7),
  heights = c(1, 1)
  ) 
```

# Fig. 2
Comparison of species composition between past and present using NMDS

## data frame for species composition
```{r}
df_comp <- function(x) {
  
  x %>% 
    dplyr::filter(value == 1) %>% 
    dplyr::select(site, time, exotic, sciname, value) %>% 
    pivot_wider(
      names_from = sciname, 
      values_from = value,
      values_fill = 0
      )
} 


df_comp_plant     <- df_comp(df_plant) 
df_comp_bird      <- df_comp(df_bird) 
df_comp_butterfly <- df_comp(df_butterfly) 
```

## NMDS | Permanova | Dispersion

Function of estimation for nmds, permanova, and dispersion
```{r}
est_nmds_prmnv <- function(x) {
  
  x %>% 
    dplyr::group_by(site, time) %>% 
    dplyr::summarise(across(where(is.numeric), sum), .groups = "drop") %>% 
    dplyr::mutate(exotic = "All") %>% 
    dplyr::select(site, time, exotic, everything()) %>% 
    dplyr::bind_rows(x) %>% 
    group_nest(exotic) %>% 
    
    # if data is plant, adding a #
    dplyr::filter(exotic != "Exotic") %>% 
    
    
    dplyr::mutate(
      
      # Rename
      exotic = recode_factor(
        exotic, 
        "All" = "All species",
        "Native" = "Native species"),
      
      # Remove columns with a total of 0
      comp = map(
        data,
        ~ dplyr::select(., where(~ is.numeric(.) && sum(.) != 0))),
      
      # Jaccard dissimilarity
      jac = map(comp, ~vegdist(., method = "jac")),
      
      # NMDS
      nmds = map(
        comp, 
        ~ metaMDS(., dist = "jac", k = 2, trace = F, trymax = 999)),
      
      # NMDS axis score
      nmds_score = map(nmds, ~scores(.) %>% pluck(1) %>% as.data.frame()),
      
      # NMDS stress value
      nmds_stress = map_dbl(nmds, ~.$stress),
      
      
      # permanova
      permanova = map(data, ~adonis2(.[,-1:-2] ~ .$time, method = "jac")),
      
      # p-value of permanova
      pval_perma = map(permanova, ~as_tibble(.) %>% dplyr::select("Pr(>F)")),
      
      
      # dispersion
      dispersion = map(data, ~betadisper(
        vegdist(.[,-1:-2], method = "jac"), .$time)
        ),
      
      # p-value of dispersion
      pval_disper = map(dispersion, ~permutest(.) %>% 
                          pluck("tab") %>% 
                          as_tibble() %>% 
                          dplyr::select("Pr(>F)")
                        )
      )
}
```

Estimate NMDS, permanova, and dispersion
```{r}
# plant
comp_plant <- est_nmds_prmnv(df_comp_plant) %>% 
  dplyr::mutate(
    exotic = recode_factor(
      exotic, 
      "All species" = "All species",
      "Native species" = "Native species")
    )

# bird
comp_bird <- est_nmds_prmnv(df_comp_bird)


# butterfly
comp_butterfly <- est_nmds_prmnv(df_comp_butterfly)
```

## plot for NMDS
Label for stress value and p-value of permanova and dispersion
```{r}
label_nmds <- function(x) {
  
  x %>% 
    dplyr::select(exotic, nmds_stress, pval_perma, pval_disper) %>% 
    unnest("pval_perma") %>% 
    na.omit() %>% 
    rename(perma = `Pr(>F)`) %>% 
    unnest("pval_disper") %>% 
    rename(disper = `Pr(>F)`) %>% 
    na.omit() %>% 
    dplyr::mutate(
      across(where(is.numeric), ~round(., digits = 3)),
      time = "Past",
      stressvalue = "Stress value:",
      permanova_cha = "Permanova: p = ",
      dispersion_cha = "Dispersion: p = "
    ) %>% 
    unite(stress, stressvalue, nmds_stress, remove = F, sep = "") %>% 
    unite(permanova, permanova_cha, perma, remove = F, sep = "") %>% 
    unite(dispersion, dispersion_cha, disper, remove = F, sep = "")
}

label_nmds_plant <- label_nmds(comp_plant)
label_nmds_bird <- label_nmds(comp_bird)
label_nmds_butterfly <- label_nmds(comp_butterfly)
```

Plot for NMDS
```{r fig.width=10, fig.height=8}
# Plant
p_nmds_plant <- comp_plant %>% 
  dplyr::select(exotic, nmds_score) %>% 
  unnest(nmds_score) %>% 
  dplyr::mutate(
    time = rep(c("Past", "Present"), 30),
    time = factor(time, levels = c("Past", "Present"))
    ) %>% 
  
  ggplot(aes(x = NMDS1, y = NMDS2, color = time)) +
  geom_point(size = 0.8) +
  stat_ellipse(
    aes(group = time, fill = time), 
    alpha = 0.1, linewidth = 0.1,
    geom = "polygon") +
  
  facet_wrap(exotic~., ncol = 1) +
  
  # Stress value
  geom_text(
    data = label_nmds_plant, aes(label = stress), 
    x = -1.8, y = 1.8, size = 1.8, color = "grey50", hjust = 0) +
  
  # P value from Permanova
  geom_text(
    data = label_nmds_plant, aes(label = permanova), 
    x = -1.8, y = 1.6, size = 1.8, color = "grey50", hjust = 0) +
  
  # P-value from Dispersion
  geom_text(
    data = label_nmds_plant, aes(label = dispersion), 
    x = -1.8, y = 1.4, size = 1.8, color = "grey50", hjust = 0) +
  
  labs(title = "Plant") + 
  scale_x_continuous(limits = c(-1.8, 1.8)) +
  scale_y_continuous(limits = c(-1.8, 1.8)) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
  theme_bw(base_size = 8) +
  theme(
    axis.title.x = element_blank(),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_text(hjust = 1),
    plot.title = element_text(hjust = 0, vjust = -5),
    
    # Legend
    legend.title = element_blank(),
    legend.position = c(0.8, 0.95),
    legend.text = element_text(size = 7),
    legend.key.height = unit(3, "mm"),
    legend.key.width = unit(3, "mm"),
    legend.background = element_blank()
    )




# Bird
p_nmds_bird <- 
  comp_bird %>% 
  dplyr::select(exotic, nmds_score) %>% 
  unnest(nmds_score) %>% 
  dplyr::mutate(
    time = c(rep(c("Present", "Past"), 7), rep(c("Past", "Present"), 7)),
    time = factor(time, levels = c("Past", "Present"))
    ) %>% 
  
  ggplot(aes(x = NMDS1, y = NMDS2, color = time, group = time)) +
  geom_point(size = 0.8) +
  stat_ellipse(
    aes(group = time, fill = time), 
    alpha = 0.1, linewidth = 0.1,
    geom = "polygon"
    ) +
  
  facet_wrap(exotic~., ncol = 1, scale = "free") +
  
  # p-value
  geom_text(
    data = label_nmds_bird, aes(label = stress), 
    x = -2, y = 1.4, hjust = 0, size = 1.8, color = "grey50") +
  
  # permanova
  geom_text(
    data = label_nmds_bird, aes(label = permanova), 
    x = -2, y = 1.25, hjust = 0, size = 1.8, color = "grey50") +
  
  # P-value from Dispersion
  geom_text(
    data = label_nmds_bird, aes(label = dispersion), 
    x = -2, y = 1.1, hjust = 0, size = 1.8, color = "grey50") +
  
  
  # adjustment
  ylim(-1.4, 1.4) +
  labs(title = "Bird", fill = "Time", color = "Time") + 
  scale_x_continuous(limits = c(-2, 2)) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
  theme_bw(base_size = 8) +
  theme(
    axis.title.y = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0, vjust = -5),
    strip.background = element_blank(),
    strip.text.x = element_text(hjust = 1),
    legend.position = "none"
  ) 





# Butterfly
p_nmds_butterfly <- comp_butterfly %>% 
  dplyr::select(exotic, nmds_score) %>% 
  unnest(nmds_score) %>% 
  dplyr::mutate(
    time = c(rep(c("Present", "Past"), 14)),
    time = factor(time, levels = c("Past", "Present"))
  ) %>% 
  
  ggplot(aes(x = NMDS1, y = NMDS2, color = time)) +
  geom_point(size = 0.8) +
  stat_ellipse(
    aes(group = time, fill = time), 
    alpha = 0.1, linewidth = 0.1,
    geom = "polygon"
    ) +
  
  facet_wrap(exotic~., ncol = 1) +
  
  # Stress value
  geom_text(
    data = label_nmds_butterfly, aes(label = stress), 
    x = -1.25, y = 1, size = 1.8, color = "grey50", hjust = 0) +
  
  # permanova
  geom_text(
    data = label_nmds_butterfly, aes(label = permanova), 
    x = -1.25, y = 0.88, size = 1.8, color = "grey50", hjust = 0) +
  
  # P-value from Dispersion
  geom_text(
    data = label_nmds_butterfly, aes(label = dispersion), 
    x = -1.25, y = 0.76, size = 1.8, color = "grey50", hjust = 0) +
  
  
  labs(title = "Butterfly") + 
  ylim(c(-1, 1)) +
  scale_x_continuous(breaks = seq(-1, 1, length = 3)) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
  theme_bw(base_size = 8) +
  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0, vjust = -5),
    strip.background = element_blank(),
    strip.text.x = element_text(hjust = 1)
    )





## Combine 
p_nmds_plant + p_nmds_bird +  p_nmds_butterfly
```

# Fig.3
Relationships between temporal changes in species richness from the past to the present and years between surveys


## Estimate species richness
```{r message=FALSE, warning=FALSE}
# Function of estimation for species richness 
est_sr <- function(x) {
  x %>% 
    dplyr::filter(value == 1 & (exotic == "Native" | exotic == "Exotic")) %>% 
    dplyr::group_by(site, time, exotic) %>% 
    dplyr::summarise(sr = n_distinct(sciname), .groups = "drop") %>% 
    pivot_wider(
      names_from = "exotic", 
      values_from = "sr",
      values_fill = list(sr = 0)
      ) %>% 
    dplyr::mutate(All = Native + Exotic) %>% 
    pivot_longer(c(All, Native, Exotic), 
                 values_to = "sr", names_to = "exotic") %>% 
    dplyr::mutate(
      exotic = recode_factor(
        exotic, 
        "All" = "All species",
        "Native" = "Native species",
        "Exotic" = "Exotic species"),
      time = recode_factor(
        time,
        "past" = "Past",
        "now"  = "Present")
      ) 
}



# Estimate a species richness
sr_plant <- est_sr(df_plant)
sr_bird <- est_sr(df_bird)
sr_butterfly <- est_sr(df_butterfly)
```

## GLM
```{r}
# Plant
glm_sr_plant <- sr_plant %>% 
  left_join(df_env, by = "site") %>% 
  pivot_wider(names_from = time, values_from = sr) %>% 
  dplyr::mutate(tem_sr = Present - Past) %>% 
  
  group_nest(exotic) %>% 
  dplyr::mutate(
    model = map(data, ~ glm(tem_sr ~ year + green_rate + area, data = .)),
    summary = map(model, ~tidy(.)),
    predict = map(model, ~ggpredict(., terms = "year"))
    )


# Bird
glm_sr_bird <- sr_bird %>% 
  left_join(df_env, by = "site") %>% 
  pivot_wider(names_from = time, values_from = sr) %>% 
  dplyr::mutate(tem_sr = Present - Past) %>% 
  group_nest(exotic) %>% 
  dplyr::mutate(
    model = map(data, ~ glm(tem_sr ~ year + green_rate + area, data = .)),
    summary = map(model, ~tidy(.))
    )


# Butterfly
glm_sr_butterfly <- sr_butterfly %>% 
  left_join(df_env, by = "site") %>% 
  pivot_wider(names_from = time, values_from = sr) %>% 
  dplyr::mutate(tem_sr = Present - Past) %>% 
  
  group_nest(exotic) %>% 
  dplyr::mutate(
    model = map(data, ~ glm(tem_sr ~ year + green_rate + area, data = .)),
    summary = map(model, ~tidy(.))
  )
```

## Table for GLM

```{r}
### Table for GLM -----
tidy_table <- function(x) {
  
  x %>% 
    dplyr::select(-data, -model) %>% 
    tidyr::unnest(summary) %>% 
    dplyr::filter(term != "(Intercept)") %>% 
    dplyr::mutate(
      # round
      across(where(is.numeric),
             ~ if_else(abs(.) <= 0.001,
                     formatC(., digits = 3, format = "e"),
                     formatC(., digits = 3, format = "fg")))
    ) %>% 
    dplyr::rename(
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
      Statistic = statistic
    ) 
}

bind_rows(
  tidy_table(glm_sr_plant) %>% dplyr::mutate(species = "Plant"),
  tidy_table(glm_sr_bird) %>% dplyr::mutate(species = "Bird"),
  tidy_table(glm_sr_butterfly) %>% dplyr::mutate(species = "Butterfly")
  ) %>% 
  dplyr::select(species, everything(), -predict) %>% 
  kableExtra::kable()
```
## Plot for GLM
```{r fig.width=8, fig.height=4}
plot_glm_tempSR_plant <- 
  ggplot() + 
  geom_point(
    data = glm_sr_plant %>% unnest(data)  %>% 
      dplyr::filter(exotic != "Exotic species"),
    aes(year, tem_sr, color = exotic, fill = exotic)
    ) +
  geom_smooth(
    data = glm_sr_plant %>% unnest(predict) %>% 
      dplyr::filter(exotic != "Exotic species"),
    aes(x, predicted, color = exotic, fill = exotic)
    ) +
  geom_ribbon(
    data = glm_sr_plant %>% unnest(predict) %>% 
      dplyr::filter(exotic != "Exotic species"),
    aes(x, ymin = conf.low, ymax = conf.high, fill = exotic),
    alpha = 0.1
    ) +
  
  scale_color_npg() +
  scale_fill_npg() +
  theme_bw(base_size = 10) +
  labs(
    title = "Plant",
    y = "Temporal changes in species richness"
    ) +
  theme(
    legend.position = c(0.7, 0.9),
    legend.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.key.size = unit(5, "mm"),
    axis.title.x = element_blank(),
    strip.placement = "outside",
    strip.background = element_blank(),
    panel.grid = element_blank()
    )



# Bird
plot_glm_tempSR_bird <- 
  ggplot() + 
  geom_point(
    data = glm_sr_bird %>% unnest(data) %>% 
      dplyr::filter(exotic != "Exotic species"),
    aes(year, tem_sr, color = exotic, fill = exotic)
  ) +
  scale_color_npg() +
  scale_fill_npg() +
  theme_bw(base_size = 10) +
  labs(
    title = "Bird",
    x = "Years between surveys") +
  theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    strip.placement = "outside",
    strip.background = element_blank(),
    panel.grid = element_blank()
    )


# Butterfly
plot_glm_tempSR_butterfly <- 
  ggplot() + 
  geom_point(
    data = glm_sr_butterfly %>% unnest(data) %>% 
      dplyr::filter(exotic != "Exotic species"),
    aes(year, tem_sr, color = exotic, fill = exotic)
  ) +
  scale_color_npg() +
  scale_fill_npg() +
  theme_bw(base_size = 10) +
  labs(title = "Butterfly") +
  theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    strip.placement = "outside",
    strip.background = element_blank(),
    panel.grid = element_blank()
  )

  



# combine
plot_glm_tempSR_plant + plot_glm_tempSR_bird + plot_glm_tempSR_butterfly
```

# Fig. 4
Relationships between temporal beta diversity (differences in species composition between the past and present) and years between surveys.

## Estimate temporal beta diversity
Function of estimate for partitioning of beta diversity
```{r message=FALSE, warning=FALSE}
# Function of estimate for partitioning of beta diversity
est_betapart <- function(x) {
  
  x %>%
    dplyr::filter(value == 1) %>% 
    dplyr::select(site, time, sciname, value) %>% 
    pivot_wider(
      names_from = sciname, 
      values_from = value,
      values_fill = 0
      ) %>%  
    dplyr::select(-time) %>% 
    dplyr::group_nest(site) %>% 
    dplyr::mutate(
      betapart = map(data, ~beta.pair(., index.family="jac")),
      tidy_beta = map(betapart, ~bind_rows(.))
    ) %>% 
    
    dplyr::select(site, tidy_beta) %>% 
    tidyr::unnest(tidy_beta) 
}
```

Estimate a temporal beta diversity
```{r}
# Plant
df_betapart_plant <- bind_rows(
  
  # all species
  est_betapart(df_plant),
  
  # native species
  df_plant %>% 
    dplyr::filter(exotic == "Native") %>% 
    est_betapart(),
  
  ) %>% 
  dplyr::mutate(
    species = "Plant",
    exotic = rep(c("All","Native"), each = 15),
    across(starts_with("beta"), ~as.numeric(.))
  ) 
  
  


# Bird
df_betapart_bird <- bind_rows(
  
  # All species
  est_betapart(df_bird),
  
  # Native species
  df_bird %>% 
    dplyr::filter(exotic == "Native") %>% 
    est_betapart()
  
  ) %>% 
  dplyr::mutate(
    species = "Bird",
    exotic = rep(c("All","Native"), each = 7),
    across(starts_with("beta"), ~as.numeric(.))
  ) 



# Butterfly
df_betapart_butterfly <- bind_rows(
  
  # All species
  est_betapart(df_butterfly),
  
  # Native species
  df_butterfly %>% 
    dplyr::filter(exotic == "Native") %>% 
    est_betapart()
  
  ) %>% 
  dplyr::mutate(
    species = "Butterfly",
    exotic = rep(c("All","Native"), each = 7),
    across(starts_with("beta"), ~as.numeric(.))
  ) 



# combine
df_betapart <- bind_rows(
  df_betapart_plant, df_betapart_bird, df_betapart_butterfly) %>% 
  pivot_longer(cols = starts_with("beta")) %>% 
  left_join(df_env, by = "site")
```

## GLM 
GLM for temporal beta diversity
```{r}
glm_betapart <- 
  df_betapart %>% 
  dplyr::group_nest(species, exotic, name) %>% 
  dplyr::mutate(
    model = map(data, ~lm(value ~ year + area + green_rate, data = .)),
    summary = map(model, ~tidy(.)),
    predict = map(model, ~ggpredict(., terms = "year")),
    
    species = fct_relevel(species, c("Plant", "Bird", "Butterfly")),
    exotic = factor(exotic, levels = c("All", "Native")),
    beta =  fct_recode(name, 
                       "Beta diversity" = "beta.jac",
                       "Turnover" = "beta.jtu",
                       "Nestedness" = "beta.jne")
    ) %>% 
  arrange(species, exotic)
```

## Table for GLM

```{r}
glm_betapart %>% 
  tidyr::unnest(summary) %>% 
  dplyr::filter(term != "(Intercept)" & name == "beta.jac") %>% 
  dplyr::select(-data, -model, -predict, -name, -beta) %>% 
  dplyr::mutate(
      # round
      across(where(is.numeric),
             ~ if_else(abs(.) <= 0.001,
                       formatC(., digits = 3, format = "e"),
                       formatC(., digits = 3, format = "fg"))),
    ) %>% 
    dplyr::rename(
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
      Statistic = statistic
      ) 
```

## Plot for GLM

```{r fig.width=8, fig.height=8}
#plot_glm_tempbeta <- 
  ggplot() + 
  geom_point(
    data = glm_betapart %>% 
      unnest(data) %>% 
      dplyr::filter(exotic != "Exotic" & name == "beta.jac"),
    aes(year, value, color = exotic, fill = exotic)
    ) +
  geom_line(
    data = glm_betapart %>% 
      dplyr::filter(
          species == "Bird" & exotic != "Exotic" & name == "beta.jac"
          ) %>% 
      unnest(predict),
    aes(x, predicted, color = exotic), linewidth = 1
    ) +
  geom_ribbon(
    data = glm_betapart %>% 
      dplyr::filter(
        species == "Bird" & exotic != "Exotic" & name == "beta.jac"
        ) %>% 
      unnest(predict),
    aes(x, ymin = conf.low, ymax = conf.high, fill = exotic),
    alpha = 0.1
    ) +
  facet_grid(.~ species, scales = "free") +
  scale_y_continuous(breaks = seq(0, 0.9, length = 4))+
  scale_color_npg(labels = c(All = "All species", Native = "Native species")) +
  scale_fill_npg(labels = c(All = "All species", Native = "Native species")) +
  labs(x = "Years between surveys", y = "Temporal beta diversity") +
  theme_bw(base_size = 10) +
  theme(
    legend.position = c(0.09, 0.9),
    legend.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 7),
    legend.key.size = unit(5, "mm"),
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text = element_text(size = 10),
    strip.text.x = element_text(hjust = 0), 
    panel.grid = element_blank()
    )
```


# Table 2
Relationships between the proportion of species lost and functional traits and environmental variables.

## Esimate lost species richness
```{r message=FALSE, warning=FALSE}
# plant
df_lost_plant <- df_plant %>% 
  pivot_wider(names_from = time, values_from = value, values_fill = 0) %>% 
  dplyr::group_by(site, sciname, growth, disper, exotic) %>% 
  dplyr::summarise(past = sum(past), now = sum(now)) %>% 
  dplyr::mutate(lost = if_else(past == 1 & now == 0, 1, 0)) %>%  
  dplyr::filter(past == 1) %>% 
  dplyr::select(-1, -now, -past) %>% 
  dplyr::left_join(df_env, by = "site")


# bird
df_lost_bird <- df_bird %>% 
  pivot_wider(names_from = time, values_from = value, values_fill = 0) %>% 
  rename(HWI = "Hand-Wing.Index",
         Beak = "Beak.Length_Culmen") %>% 
  dplyr::group_by(site, sciname, exotic) %>% 
  dplyr::summarise(
    past = sum(past), now = sum(now),
    Beak = mean(Beak), HWI = mean(HWI), Mass = mean(Mass)
    ) %>% 
  dplyr::mutate(lost = if_else(past == 1 & now == 0, 1, 0)) %>% 
  dplyr::filter(past == 1) %>% 
  dplyr::select(-now, -past) %>% 
  dplyr::left_join(df_env, by = "site") %>% 
  ungroup()


# butterfly
df_lost_butterfly <- df_butterfly %>% 
  pivot_wider(names_from = time, values_from = value, values_fill = 0) %>% 
  dplyr::group_by(site, sciname, exotic) %>% 
  dplyr::summarise(
    past = sum(past), now = sum(now),
    FlightDuration = mean(FlightDuration), WS = mean(WS), 
    hostplant = mean(hostplant)
    ) %>% 
  dplyr::mutate(lost = if_else(past == 1 & now == 0, 1, 0)) %>% 
  dplyr::filter(past == 1) %>% 
  dplyr::select(-now, -past) %>% 
  dplyr::left_join(df_env, by = "site") %>% 
  ungroup()
```

## GLM
GLM for proportion of lost species (only for native species)
```{r warning=FALSE}
# plant
glm_lost_plant <- 
  df_lost_plant %>% 
  dplyr::filter(exotic == "Native") %>% 
  glm(lost ~ year + area + green_rate + growth + disper + 
        year:growth + year:disper,
      family = "binomial", data = .)


# bird
glm_lost_bird <- df_lost_bird %>% 
  dplyr::filter(exotic == "Native") %>% 
  glm(lost ~ year + area + green_rate + Mass + Beak + HWI +
        year:Mass + year:Beak + year:HWI,
      family = "binomial", data = .) 


# butterfly
glm_lost_butterfly <- df_lost_butterfly %>% 
  glm(lost ~ year + area + green_rate + FlightDuration + WS + hostplant + 
        year:FlightDuration + year:WS + year:hostplant,
      family = "binomial", data = .) 
```
## Table for GLM
```{r}
# function for table of GLM 
tidy_table_lost <- function(x) {
  
  x %>% 
    tidy() %>%   
    dplyr::filter(term != "(Intercept)") %>% 
    dplyr::mutate(
      # round
      across(where(is.numeric),
             ~ if_else(abs(.) <= 0.001,
                       formatC(., digits = 3, format = "e"),
                       formatC(., digits = 3, format = "fg")))
    ) %>% 
    dplyr::rename(
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
      Statistic = statistic
    ) 
}



bind_rows(
  tidy_table_lost(glm_lost_plant) %>% dplyr::mutate(species = "Plant"),
  tidy_table_lost(glm_lost_bird) %>% dplyr::mutate(species = "Bird"),
  tidy_table_lost(glm_lost_butterfly) %>% dplyr::mutate(species = "Butterfly")
  ) %>% 
  dplyr::select(species, "Explanatory variables", Estimate, p.value) %>% 
  kableExtra::kable()
```


