---
title: "Bird"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  root.dir = "/temp")
```

# Package

```{r package, include=FALSE}
pacman::p_load(
  tidyverse,  # data management
  rtry, austraits,  # trait
  lme4, lmerTest, broom,  # GLM
  ggeffects, patchwork, ggsci,  # ggplot
  vegan, # composition
  ggcorrplot, ggbiplot
  )
```

# Data frame

Load a data frame

```{r , include=FALSE}
# survey data
bird <- read_csv("data/bird_survey.csv")

# scientific name
bird_sciname <- read_csv("data/bird_sciname.csv")

# trait 
#bird_trait <- read_csv("data/bird_trait_avonet.csv")
bird_trait_life   <- read_csv("data/bird_trait_birdlife.csv")
bird_trait_ebird  <- read_csv("data/bird_trait_ebird.csv")
bird_trait_tree   <- read_csv("data/bird_trait_birdtree.csv")
bird_trait_avonet <- read_csv("data/bird_trait_avonet.csv")

# exotic
bird_exotic <- read_csv("data/bird_exotic.csv")

# Environmental variables
env <- read_csv("data/Environment.csv")
```

Data frame for bird trait
```{r}
df_bird_trait <- 
  bind_rows(
  
    # birdlife
    bird_trait_life %>% 
      dplyr::select(Species1, Beak.Length_Culmen:Mass) %>% 
      dplyr::rename(sciname = Species1),
    
    # ebird
    bird_trait_ebird %>% 
      dplyr::select(Species2, Beak.Length_Culmen:Mass) %>% 
      dplyr::rename(sciname = Species2),
    
    # birdtree
    bird_trait_tree%>% 
      dplyr::select(Species3, Beak.Length_Culmen:Mass) %>% 
      dplyr::rename(sciname = Species3),
    
    # Avonet
    bird_trait_avonet %>% 
      dplyr::select(Species1_BirdLife, Beak.Length_Culmen:Tail.Length) %>% 
      mutate(Mass = NA) %>% 
      dplyr::rename(sciname = Species1_BirdLife)
    ) %>% 
  group_by(sciname) %>% 
  dplyr::summarise(across(where(is.numeric), ~mean(., na.rm = TRUE))) %>% 
  print()
```
Combine the all data
```{r }
df_bird <- 
  bird %>% 
    as_tibble() %>% 
    mutate(time = if_else(str_detect(site, "過去"), "past", "now")) %>% 
    select(site, time, everything()) %>% 
    mutate(site = str_sub(site, end = -3)) %>% 
    pivot_longer(-c(site, time), names_to = "species") %>% 
    filter(value == 1) %>% 
    
    # combine scientific name + Exotic 
    left_join(
      bird_sciname %>% 
        select(species, sciname) %>% 
        mutate(exotic = "Native") %>% 
        filter(species != "カササギ") %>% 
        bind_rows(bird_exotic), 
      by = "species") %>% 
    
    
    # combine trait data
    left_join(df_bird_trait, by = "sciname") %>% 
    
    print()
```
```{r}
# Number of species of no trait data
df_bird %>% 
  filter(is.na(Mass)) %>% 
  distinct(species, sciname)



#
df_bird %>% 
  na.omit() %>% 
  select(where(is.numeric) & -value) %>% 
  prcomp(scale = TRUE) %>% 
  biplot()
```



# Species richness

Estimate a species richness

```{r}
sr_bird <-
  df_bird %>% 
    group_by(site, time, exotic) %>% 
    dplyr::summarise(sr = n_distinct(species)) %>% 
    filter(exotic != "unknown") %>% 
    pivot_wider(names_from = "exotic", values_from = "sr",
                values_fill = list(sr = 0)) %>% 
    mutate(All = Native + Exotic) %>% 
    pivot_longer(c(All, Native, Exotic), 
                 values_to = "sr", names_to = "exotic") %>% 
    mutate(
    exotic = recode_factor(
      exotic, 
      "All" = "All species",
      "Native" = "Native species",
      "Exotic" = "Exotic species"),
    time = recode_factor(
      time,
      "past" = "Past",
      "now"  = "Now")
    ) %>% 
    print()
```

### Boxplot

Comparison of species richness

```{r}
box_sr_bird <- 
  ggplot(data = sr_bird,
         aes(x = time, y = sr, fill = time)) +
    geom_boxplot() +
    facet_wrap(~exotic, scales = "free", ncol = 3) +
    labs(
      x = "Time",
      y = "Species richness",
      fill = "Time"
    ) +
    theme_bw(base_size = 8) +
    theme(
      panel.grid = element_blank(),
      legend.position = "none"
      )

box_sr_bird
```

```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  box_sr_bird,
  file = "output/box_sr_bird.png",
  width = 160, height = 60, units = "mm", dpi = 500)
```



# Lost and add species

Estimate a lost and added species richness

```{r}
df_bird_lost_add <-
  df_bird %>% 
    filter(exotic != "unknown") %>% 
    pivot_wider(names_from = time, values_from = value) %>% 
    dplyr::rename(HWI = "Hand-Wing.Index") %>% 
    mutate(
      now = if_else(is.na(now), 0, 1),
      past = if_else(is.na(past), 0, 1),
      lost = if_else(past == 1 & now == 0, 1, 0),
      add  = if_else(past == 0 & now == 1, 1, 0)
      ) %>% 
    left_join(env, by = "site") %>% 
    mutate(exotic = factor(exotic, levels = c("Native", "Exotic"))) %>% 
    print()
```

### Lost species

#### Trait
Relationships between lost of species and trait

GLM
```{r}
glm_lost_trait_bird <- 
  df_bird_lost_add %>% 
    group_by(exotic) %>% 
    nest() %>% 
    mutate(
       fit = map(data, ~glm(lost ~ Mass + Beak.Length_Culmen + HWI, 
                            family = "binomial",  data = .)),
       summary = map(fit, ~tidy(.)),
       pred = map(fit, ~ggpredict(., terms="HWI [all]"))
       ) %>% 
    ungroup() %>% 
    print()
```
Table for result of GLM
```{r}
tb_glm_lost_trait_bird <- 
  glm_lost_trait_bird %>% 
    dplyr::select(exotic, summary) %>% 
    unnest(summary) %>% 
    filter(term != "(Intercept)") %>% 
    mutate(
      across(where(is.numeric),
             ~if_else(. < 0.001,
                      formatC(., digits = 3, format = "fg"),
                      formatC(., digits = 3, format = "e"))
      )) %>% 
    dplyr::rename(
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
      Statistic = statistic) %>% 
    print()
```
Plot for the result of GLM
```{r}
theme_set(theme_classic(base_size = 9))

# Hand-Wing.Index
p_glm_lost_bird_hwi <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = HWI, y = lost, color = exotic, shape = exotic)) +
    geom_smooth(
      data = glm_lost_trait_bird %>% 
        select(exotic, pred) %>% 
        unnest(pred) %>% 
        filter(exotic == "Native"),
      aes(x = x, y = predicted, color = exotic),
      method = "glm", method.args=list(family="binomial"), se = FALSE) +
    labs(
      x = "Hand-Wing.Index",
      y = "Probability of species loss"
      ) +
      scale_color_simpsons() +
      theme(
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 9),
        legend.position = c(0.3, 0.8),
        legend.title = element_blank()
        )

# Body mass
p_glm_lost_bird_mass <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = Mass, y = lost, color = exotic, shape = exotic)) +
    labs(
      x = "Body mass",
      y = ""
      ) +
    scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )


# Beak length
p_glm_lost_bird_beak <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = Beak.Length_Culmen, y = lost, color = exotic, shape = exotic)) +
      labs(
        x = "Beak length",
        y = ""
      ) +
      scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )



# Combine
p_glm_lost_bird_hwi + p_glm_lost_bird_mass + p_glm_lost_bird_beak 
```
```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  p_glm_lost_bird_hwi + p_glm_lost_bird_mass + p_glm_lost_bird_beak ,
  file = "output/plot_glm_lost_trait_bird.png",
  width = 160, height = 90, units = "mm", dpi = 500)
```


#### Environment
Relationships between lost of species and trait using GLM

GLM
```{r}
glm_lost_env_bird <- 
  df_bird_lost_add %>% 
    group_by(exotic) %>% 
    nest() %>% 
    mutate(
       fit = map(data, ~glm(lost ~ year + area + green_rate, 
                            family = "binomial",  data = .)),
       summary = map(fit, ~tidy(.)),
       pred = map(fit, ~ggpredict(.))
       ) %>% 
    ungroup() %>% 
    print()
```
Table for result of GLM
```{r}
tb_glm_lost_env_bird <- 
  glm_lost_env_bird %>% 
    dplyr::select(exotic, summary) %>% 
    unnest(summary) %>% 
    filter(term != "(Intercept)") %>% 
    dplyr::mutate(
      across(where(is.numeric),
             ~if_else(abs(.) >= 0.001,
                      round(., digits = 3), .)
      )) %>% 
    dplyr::rename(
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
      Statistic = statistic) %>% 
    print()
```
Plot for the result of GLM
```{r}
theme_set(theme_classic(base_size = 9))

# Year
p_glm_lost_bird_year <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = year, y = lost, color = exotic, shape = exotic)) +
    geom_smooth(
      data = glm_lost_env_bird %>% 
        select(exotic, pred) %>% 
        unnest(pred) %>% 
        unnest(pred) %>% 
        filter(exotic == "Native" & group == "year"),
      aes(x = x, y = predicted, color = exotic),
      method = "glm", method.args=list(family="binomial"), se = FALSE) +
    labs(
      x = "Year",
      y = "Probability of species loss"
      ) +
      scale_color_simpsons() +
      theme(
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 9),
        legend.position = c(0.3, 0.8),
        legend.title = element_blank()
        )

# Area
p_glm_lost_bird_area <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = area/10000, y = lost, color = exotic, shape = exotic)) +
    geom_smooth(
      data = glm_lost_env_bird %>% 
        select(exotic, pred) %>% 
        unnest(pred) %>% 
        unnest(pred) %>% 
        filter(exotic == "Native" & group == "area"),
      aes(x/10000, predicted, color = exotic),
      method = "glm", method.args=list(family="binomial"), se = FALSE) +
    labs(
      x = "Area",
      y = ""
      ) +
    scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )


# Green rate
p_glm_lost_bird_green <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = green_rate, y = lost, color = exotic, shape = exotic)) +
      labs(
        x = "Green rate",
        y = ""
      ) +
      scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )



# Combine
p_glm_lost_bird_year + p_glm_lost_bird_area + p_glm_lost_bird_green
```
```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  p_glm_lost_bird_year + p_glm_lost_bird_area + p_glm_lost_bird_green,
  file = "output/plot_glm_lost_env_bird.png",
  width = 160, height = 90, units = "mm", dpi = 500)
```


### Added species

#### Trait
Relationships between species immigration and trait

GLM
```{r}
glm_add_trait_bird <- 
  df_bird_lost_add %>% 
    group_by(exotic) %>% 
    nest() %>% 
    mutate(
       fit = map(data, ~glm(add ~ Mass + Beak.Length_Culmen + HWI, 
                            family = "binomial",  data = .)),
       summary = map(fit, ~tidy(.)),
       pred = map(fit, ~ggpredict(.))
       ) %>% 
    ungroup() %>% 
    print()
```
Table for result of GLM
```{r}
tb_glm_add_trait_bird <- 
  glm_add_trait_bird %>% 
    dplyr::select(exotic, summary) %>% 
    unnest(summary) %>% 
    filter(term != "(Intercept)") %>% 
    mutate(
      across(where(is.numeric),
             ~if_else(. < 0.001,
                      formatC(., digits = 3, format = "fg"),
                      formatC(., digits = 3, format = "e"))
      )) %>% 
    dplyr::rename(
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
      Statistic = statistic) %>% 
    print()
```
Plot for the result of GLM
```{r}
theme_set(theme_classic(base_size = 9))

# Hand-Wing.Index
p_glm_add_bird_hwi <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = HWI, y = add, color = exotic, shape = exotic)) +
    labs(
      x = "Hand-Wing.Index",
      y = "Probability of species immigration"
      ) +
      scale_color_simpsons() +
      theme(
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 9),
        legend.position = c(0.3, 0.8),
        legend.title = element_blank()
        )

# Body mass
p_glm_add_bird_mass <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = Mass, y = add, color = exotic, shape = exotic)) +
    labs(
      x = "Body mass",
      y = ""
      ) +
    scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )


# Beak length
p_glm_add_bird_beak <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = Beak.Length_Culmen, y = add, color = exotic, shape = exotic)) +
      labs(
        x = "Beak length",
        y = ""
      ) +
      scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )



# Combine
p_glm_add_bird_hwi + p_glm_add_bird_mass + p_glm_add_bird_beak 
```
```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  p_glm_add_bird_hwi + p_glm_add_bird_mass + p_glm_add_bird_beak,
  file = "output/plot_glm_add_trait_bird.png",
  width = 160, height = 90, units = "mm", dpi = 500)
```





#### Environment
Relationships between immigration species and trait

GLM
```{r}
glm_add_env_bird <- 
  df_bird_lost_add %>% 
    group_by(exotic) %>% 
    nest() %>% 
    mutate(
       fit = map(data, ~glm(add ~ year + area + green_rate, 
                            family = "binomial",  data = .)),
       summary = map(fit, ~tidy(.)),
       pred = map(fit, ~ggpredict(.))
       ) %>% 
    ungroup() %>% 
    print()
```
Table for result of GLM
```{r}
tb_glm_add_env_bird <- 
  glm_add_env_bird %>% 
    dplyr::select(exotic, summary) %>% 
    unnest(summary) %>% 
    filter(term != "(Intercept)") %>% 
    dplyr::mutate(
      across(where(is.numeric),
             ~if_else(abs(.) >= 0.001,
                      round(., digits = 3), .)
      )) %>% 
    dplyr::rename(
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
      Statistic = statistic) %>% 
    print()
```
Plot for the result of GLM
```{r}
theme_set(theme_classic(base_size = 9))

# Year
p_glm_add_bird_year <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = year, y = add, color = exotic, shape = exotic)) +
    geom_smooth(
      data = glm_add_env_bird %>% 
        select(exotic, pred) %>% 
        unnest(pred) %>% 
        unnest(pred) %>% 
        filter(exotic == "Native" & group == "year"),
      aes(x = x, y = predicted, color = exotic),
      method = "glm", method.args=list(family="binomial"), se = FALSE) +
    labs(
      x = "Year",
      y = "Probability of species immigration"
      ) +
      scale_color_simpsons() +
      theme(
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 9),
        legend.position = c(0.3, 0.8),
        legend.title = element_blank()
        )

# Area
p_glm_add_bird_area <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = area/10000, y = add, color = exotic, shape = exotic)) +
    geom_smooth(
      data = glm_add_env_bird %>% 
        select(exotic, pred) %>% 
        unnest(pred) %>% 
        unnest(pred) %>% 
        filter(exotic == "Native" & group == "area"),
      aes(x/10000, predicted, color = exotic),
      method = "glm", method.args=list(family="binomial"), se = FALSE) +
    labs(
      x = "Area",
      y = ""
      ) +
    scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )


# Green rate
p_glm_add_bird_green <- 
  ggplot() +
    geom_point(
      data = df_bird_lost_add,
      aes(x = green_rate, y = add, color = exotic, shape = exotic)) +
      labs(
        x = "Green rate",
        y = ""
      ) +
      scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )



# Combine
p_glm_add_bird_year + p_glm_add_bird_area + p_glm_add_bird_green
```

```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  p_glm_add_bird_year + p_glm_add_bird_area + p_glm_add_bird_green,
  file = "output/plot_glm_add_env_bird.png",
  width = 160, height = 90, units = "mm", dpi = 500)
```

# Species composition
```{r}
df_bird %>% 
  dplyr::select(site:exotic, -sciname) %>% 
  pivot_wider(names_from = "species", values_from = value, 
              values_fill = 0) %>% 
  filter(exotic == "Native") %>% {
    
    dplyr::select(., 1:3) ->> row_bird
    dplyr::select(., -1:-3) ->> com_bird
  }
```
## Permanova
```{r}
permanova_bird <- 
  adonis2(com_bird ~ row_bird$time, 
          method = "jac", permutations = 999) %>% 
  print()

permanova_bird$`Pr(>F)`[1]
```
## NMDS
```{r}
nmds_bird <-
  metaMDS(com_bird, distance = "jac", k = 2, trace = FALSE,
          trymax = 999) 

nmds_bird
```

Plot for result of nmds
```{r}
p_nmds_bird <- 
  nmds_bird %>% 
  scores() %>% 
  pluck(1) %>% 
  bind_cols(row_bird) %>% 
  mutate(time = recode_factor(time, "past" = "Past", "now" = "Now")) %>% 
    ggplot(aes(x = NMDS1, y = NMDS2, color = time, group = time)) +
      geom_point() +
      stat_ellipse(
        aes(group = time, fill = time), 
        alpha = 0.1,
        geom = "polygon") +
  
    # p-value
    annotate(
      "text", x = -1.5, y = 1.2, hjust = 0, size = 2, color = "grey50",
      label = paste0("Stress value : ", format(nmds_bird$stress, digits = 3))) +
  
    # permanova
    annotate(
      "text", x = -1.5, y = 1.1, hjust = 0, size = 2, color = "grey50",
      label = paste0("Permanova : p=", permanova_bird$`Pr(>F)`[1])) +
    theme(
      legend.position = c(0.9, 0.9),
      legend.title = element_blank()
      )

p_nmds_bird
```
```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  p_nmds_bird,
  file = "output/plot_nmds_bird.png",
  width = 90, height = 90, units = "mm", dpi = 500)
```










