---
title: "Bird"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  root.dir = "/temp")
```

# Package

```{r package, include=FALSE}
pacman::p_load(
  tidyverse,  # data management
  rtry, austraits,  # trait
  lme4, lmerTest, broom,  # GLM
  ggeffects, patchwork, ggsci,  # ggplot
  vegan, # composition
  ggcorrplot, ggbiplot
  )
```

# Data frame

Load a data frame

```{r , include=FALSE}
# survey data
bird <- read_csv("data/bird_survey.csv")

# scientific name
bird_sciname <- read_csv("data/bird_sciname.csv")

# trait 
#bird_trait <- read_csv("data/bird_trait_avonet.csv")
bird_trait_life   <- read_csv("data/bird_trait_birdlife.csv")
bird_trait_ebird  <- read_csv("data/bird_trait_ebird.csv")
bird_trait_tree   <- read_csv("data/bird_trait_birdtree.csv")
bird_trait_avonet <- read_csv("data/bird_trait_avonet.csv")

# exotic
bird_exotic <- read_csv("data/bird_exotic.csv")

# Environmental variables
env <- read_csv("data/Environment.csv")
```

Data frame for bird trait
```{r}
df_bird_trait <- 
  bind_rows(
  
    # birdlife
    bird_trait_life %>% 
      dplyr::select(Species1, Beak.Length_Culmen:Mass) %>% 
      dplyr::rename(sciname = Species1),
    
    # ebird
    bird_trait_ebird %>% 
      dplyr::select(Species2, Beak.Length_Culmen:Mass) %>% 
      dplyr::rename(sciname = Species2),
    
    # birdtree
    bird_trait_tree%>% 
      dplyr::select(Species3, Beak.Length_Culmen:Mass) %>% 
      dplyr::rename(sciname = Species3),
    
    # Avonet
    bird_trait_avonet %>% 
      dplyr::select(Species1_BirdLife, Beak.Length_Culmen:Tail.Length) %>% 
      mutate(Mass = NA) %>% 
      dplyr::rename(sciname = Species1_BirdLife)
    ) %>% 
  group_by(sciname) %>% 
  dplyr::summarise(across(where(is.numeric), ~mean(., na.rm = TRUE))) %>% 
  print()
```
Combine the all data
```{r }
df_bird <- 
  bird %>% 
    as_tibble() %>% 
    dplyr::mutate(time = if_else(str_detect(site, "過去"), "past", "now")) %>% 
    dplyr::select(site, time, everything()) %>% 
    dplyr::mutate(site = str_sub(site, end = -3)) %>% 
    pivot_longer(-c(site, time), names_to = "species") %>% 
    dplyr::filter(value == 1) %>% 
    
    # combine scientific name + Exotic 
    left_join(
      bird_sciname %>% 
        dplyr::select(species, sciname) %>% 
        dplyr::mutate(exotic = "Native") %>% 
        dplyr::filter(species != "カササギ") %>% 
        bind_rows(bird_exotic), 
      by = "species") %>% 
    
    
    # combine trait data
    left_join(df_bird_trait, by = "sciname") %>% 
    
    print()
```
```{r}
# Number of species of no trait data
df_bird %>% 
  filter(is.na(Mass)) %>% 
  distinct(species, sciname)



# PCA
df_bird %>% 
  na.omit() %>% 
  dplyr::select(where(is.numeric) & -value) %>% 
  prcomp(scale = TRUE) %>% 
  biplot()
```

# Analysis

## Species richness

Estimate a species richness

```{r}
sr_bird <-
  df_bird %>% 
    group_by(site, time, exotic) %>% 
    dplyr::summarise(sr = n_distinct(species)) %>% 
    filter(exotic != "unknown") %>% 
    pivot_wider(names_from = "exotic", values_from = "sr",
                values_fill = list(sr = 0)) %>% 
    dplyr::mutate(All = Native + Exotic) %>% 
    pivot_longer(c(All, Native, Exotic), 
                 values_to = "sr", names_to = "exotic") %>% 
    dplyr::mutate(
      exotic = recode_factor(
        exotic, 
        "All" = "All species",
        "Native" = "Native species",
        "Exotic" = "Exotic species"),
      time = recode_factor(
        time,
        "past" = "Past",
        "now"  = "Present")
      ) %>% 
    ungroup() %>% 
    print()
```

Comparison between present and past using Wilcoxon signed rank test

```{r}
wlcx_sr_bird_time <-
  sr_bird %>% 
  nest_by(exotic) %>% 
  mutate(
    wlcx_test = map(data, ~wilcox.test(sr ~ time, data =., paired = TRUE)),
    wlcx_summary = map(wlcx_test, ~tidy(.))
    )
```

P-value of wilcoxon test in the boxplot 
```{r}
p_val_wlcx_sr_bird_time <- 
  wlcx_sr_bird_time %>% 
   unnest(wlcx_summary) %>% 
   dplyr::select(exotic, p.value) %>% 
   dplyr::mutate(
     char = "p = ",
     p.value = round(p.value, 3)
     ) %>% 
   unite(col = char_pval, char, p.value, remove = F, sep = "") %>% 
   print()  
```



Comparison of species richness

```{r}
box_sr_bird <- 
  ggplot(data = sr_bird, aes(x = time, y = sr)) +
    geom_boxplot(aes(fill = time)) + 
    geom_point(color = "grey50", size = 0.5, alpha = 0.5) +
    geom_line(aes(group = interaction(site, exotic)), 
              color = "grey50", linewidth = 0.3, alpha = 0.5) +
    facet_wrap(~exotic, ncol = 3) +
    geom_signif(
      data = p_val_wlcx_sr_bird_time,
      aes(y_position = c(148, 140, 20), 
          xmin = c(1, 1, 1), xmax = c(2, 2, 2),
          annotations = char_pval),
          size = 0.2, textsize = 2, manual = TRUE) +
    scale_y_continuous(limits = c(0, 160)) +
    labs(
      x = "Time",
      y = "Species richness"
    ) +
    theme_bw(base_size = 8) +
    theme(
      panel.grid = element_blank(),
      legend.position = "none",
      strip.background = element_blank(),
      strip.text = element_text(size = 8)
      )

box_sr_bird
```

```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  box_sr_bird,
  file = "output/box_sr_bird.png",
  width = 160, height = 60, units = "mm", dpi = 500)
```



## Lost species

Estimate a number of lost species

```{r}
df_bird_lost <-
  df_bird %>% 
    pivot_wider(names_from = time, values_from = value) %>% 
    dplyr::rename(HWI = "Hand-Wing.Index") %>% 
    dplyr::mutate(
      past = if_else(is.na(past), 0, 1),
      lost = if_else(past == 1 & is.na(now), 1, 0)
      ) %>% 
    left_join(env, by = "site") %>% 
    dplyr::filter(exotic != "unknown" & past == 1) %>% 
    dplyr::mutate(exotic = factor(exotic, levels = c("Native", "Exotic"))) 
```


### GLM
Relationships between lost of species and trait

GLM
```{r}
glm_bird_lost <- 
  df_bird_lost %>% 
    dplyr::filter(exotic == "Native") %>% 
    glm(lost ~ Mass + Beak.Length_Culmen + HWI + year + area + green_rate, 
        family = "binomial",  data = .)

summary(glm_bird_lost)
```
Table for result of GLM
```{r}
tb_glm_lost_bird <- 
  glm_lost_bird %>% 
    tidy() %>% 
    dplyr::filter(term != "(Intercept)") %>% 
    dplyr::mutate(
      across(where(is.numeric),
             ~if_else(. < 0.001,
                      formatC(., digits = 3, format = "e"),
                      formatC(., digits = 3, format = "fg"))
      )) %>% 
    dplyr::rename(
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
      Statistic = statistic) %>% 
    print()
```

```{r eval=FALSE, include=FALSE}
write.csv(tb_glm_lost_bird, "output/glm_lost_bird.csv")
```

## Colonized species

Estimate a number of colonized species

```{r}
df_bird_colo <-
  df_bird %>% 
    pivot_wider(names_from = time, values_from = value) %>% 
    dplyr::rename(HWI = "Hand-Wing.Index") %>% 
    dplyr::mutate(
      colo = if_else(is.na(past) & now == 1, 1, 0)
      ) %>% 
    left_join(env, by = "site") %>% 
    dplyr::filter(exotic != "unknown" & now == 1) %>% 
    dplyr::mutate(exotic = factor(exotic, levels = c("Native", "Exotic"))) 
```


### GLM
```{r}
glm_bird_colo <- 
  df_bird_colo %>% 
    dplyr::filter(exotic == "Native") %>% 
    glm(colo ~ Mass + Beak.Length_Culmen + HWI + year + area + green_rate,
        family = "binomial",  data = .)

summary(glm_bird_colo)
```
Table for result of GLM
```{r}
tb_glm_bitd_colo <- 
  glm_bird_colo %>% 
    tidy() %>% 
    dplyr::filter(term != "(Intercept)") %>% 
    dplyr::mutate(
      across(where(is.numeric),
             ~if_else(. < 0.001,
                      formatC(., digits = 3, format = "e"),
                      formatC(., digits = 3, format = "fg"))
      )) %>% 
    dplyr::rename(
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
      Statistic = statistic) 
```

```{r eval=FALSE, include=FALSE}
write.csv(tb_glm_bitd_colo, "output/glm_colo_bird.csv")
```



# Species composition
```{r}
df_bird %>% 
  dplyr::select(site:exotic, -sciname) %>% 
  pivot_wider(names_from = "species", values_from = value, values_fill = 0) %>% 
  filter(exotic == "Native") %>% {
    
    dplyr::select(., 1:3) ->> row_bird
    dplyr::select(., -1:-3) ->> com_bird
  }
```
## Permanova
```{r}
permanova_bird <- 
  adonis2(com_bird ~ row_bird$time, 
          method = "jac", permutations = 999) %>% 
  print()

permanova_bird$`Pr(>F)`[1]
```
## NMDS
```{r}
nmds_bird <-
  metaMDS(com_bird, distance = "jac", k = 2, trace = FALSE,
          trymax = 999) 

nmds_bird
```

Plot for result of nmds
```{r}
p_nmds_bird <- 
  nmds_bird %>% 
  scores() %>% 
  pluck(1) %>% 
  bind_cols(row_bird) %>% 
  dplyr::mutate(time = recode_factor(time, 
                                     "past" = "Past", 
                                     "now" = "Present")
                ) %>% 
    ggplot(aes(x = NMDS1, y = NMDS2, color = time, group = time)) +
      geom_point() +
      stat_ellipse(
        aes(group = time, fill = time), 
        alpha = 0.1,
        geom = "polygon") +
  
    # p-value
    annotate(
      "text", x = 0.6, y = -1.3, hjust = 0, size = 2, color = "grey50",
      label = paste0("Stress value : ", format(nmds_bird$stress, digits = 3))) +
  
    # permanova
    annotate(
      "text", x = 0.6, y = -1.4, hjust = 0, size = 2, color = "grey50",
      label = paste0("Permanova : p=", permanova_bird$`Pr(>F)`[1])) +
  
    theme_classic(base_size = 8) +
    theme(
      legend.position = c(0.15, 0.9),
      legend.title = element_blank()
      ) 
  
p_nmds_bird
```
```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  p_nmds_bird,
  file = "output/plot_nmds_bird.png",
  width = 90, height = 90, units = "mm", dpi = 500)
```










