---
title: "Plant"
author: "ubd2751"
date: "2023-04-04"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  root.dir = "/temp")
```

# Analysis

Number of species without dispersal type data

```{r}
df_plant %>% 
  distinct(species, .keep_all = TRUE) %>% 
  dplyr::filter(disper == "water" & time == "past") 
```




## Lost species

Estimate a number of lost species

```{r}
df_plant_lost <- df_plant %>% 
  pivot_wider(names_from = time, values_from = value) %>% 
  dplyr::mutate(lost = if_else(past == 1 & now == 0, 1, 0)) %>% 
  dplyr::left_join(env, by = "site") %>% 
  dplyr::filter(!is.na(exotic) & past == 1) %>% 
  dplyr::mutate(across(where(is.character), as.factor))
```

Confirmation of number of lost species across dispersal type
```{r}
df_plant_lost %>% 
  dplyr::group_by(exotic, disper) %>% 
  dplyr::summarise(n = sum(lost)/n()) %>% 
  na.omit()
```

### GLM

Relationships between lost species and trait, environmental variables

```{r}
glm_plant_lost <- df_plant_lost %>% 
  dplyr::mutate(across(where(is.character), as.factor)) %>% 
  group_nest(exotic) %>% 
  dplyr::mutate(
    
    # GLM
    fit = map(
      data, 
      ~glm(lost ~ year + area + green_rate + growth + disper, 
           family = "binomial", data = .)),
      summary_fit = map(fit, ~tidy(.)),
      
      
      # Growth form
      multi_grow = map(fit, ~glht(., linfct = mcp(growth = "Tukey"))),
      multi_grow_conf = map(multi_grow, ~confint(.) %>% tidy()),
      multi_grow_summary = map(multi_grow, ~tidy(.)),
      multi_grow_char = map(multi_grow, ~cld(., decreasing = F)),
      
      
      # Dispersal type
      multi_disper = map(fit, ~glht(., linfct = mcp(disper = "Tukey"))),
      multi_disper_conf = map(multi_disper, ~confint(.) %>% tidy()),
      multi_dipser_summary = map(multi_disper, ~tidy(.)),
      multi_disper_char = map(multi_disper, ~cld(., decreasing = F)),
      
      
      # Environment
      pred_year = map(fit, ~ggpredict(., terms = "year")),
      pred_area = map(fit, ~ggpredict(., terms = "area")),
      pred_green = map(fit, ~ggpredict(., terms = "green_rate"))
      ) 
```

Partial r squared
```{r eval=FALSE, include=FALSE}

# All species
glm_lost_plant <- 
  glm(lost ~ year + area + green_rate + growth + disper, 
      family = "binomial", data = df_plant_lost)

rsq_glm_lost_plant <- rsq.partial(glm_lost_plant, adj = TRUE)



# Native species
df_plant_lost_native <-
  df_plant_lost %>% 
    dplyr::filter(exotic == "Native" & !is.na(growth) & !is.na(disper)) 

glm_lost_plant_native <- 
  glm(lost ~ year + area + green_rate + growth + disper, 
      family = "binomial", data = df_plant_lost_native)

rsq_glm_lost_plant_native <- rsq.partial(glm_lost_plant_native, adj = TRUE)




# Exotic species
df_plant_lost_exotic <-
  df_plant_lost %>% 
    dplyr::filter(exotic == "Exotic" & !is.na(growth) & !is.na(disper)) 

glm_lost_plant_exotic <- 
  glm(lost ~ year + area + green_rate + growth + disper, 
      family = "binomial", data = df_plant_lost_exotic)

rsq_glm_lost_plant_exotic <- rsq.partial(glm_lost_plant_exotic, adj = TRUE)


df_rsq_glm_lost_plant <- 
  bind_rows(
    rsq_glm_lost_plant[2:3],
    rsq_glm_lost_plant_native[2:3],
    rsq_glm_lost_plant_exotic[2:3]
    ) %>% 
    dplyr::mutate(Nativity = rep(c("All", "Native","Exotic"), each = 5)) %>% 
  print()
```
```{r eval=FALSE, include=FALSE}
write.csv(df_rsq_glm_lost_plant, "output/rsq_glm_lost_plant.csv")
```



Table for result of GLM

```{r}
tb_glm_plant_lost <- 
  glm_plant_lost %>% 
    dplyr::select(exotic, summary_fit) %>% 
    unnest(summary_fit) %>% 
    dplyr::filter(term != "(Intercept)") %>% 
    dplyr::mutate(
      across(where(is.numeric),
             ~ifelse(. > -0.001 & . < 0.001,
                      formatC(., digits = 3, format = "e"),
                      formatC(., digits = 3, format = "fg"))
      )) %>% 
    dplyr::rename(
      Nativity = exotic,
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
      Statistic = statistic) %>% 
    print()
```

```{r eval=FALSE, include=FALSE}
write.csv(tb_glm_plant_lost, "output/glm_lost_plant.csv")
```



### Plot

#### Growth form
```{r}
plot_glm_pant_lost_growth <- glm_plant_lost %>% 
  dplyr::select(exotic, multi_grow_conf) %>% 
  unnest(multi_grow_conf) %>% 
  dplyr::mutate(
    exotic = recode_factor(
      exotic,
      "Native" = "Native spcies",
      "Exotic" = "Exotic species"
      )
    ) %>% 
  
  ggplot(aes(x = estimate, y = contrast, xmin = conf.low, xmax = conf.high)) +
    geom_errorbar(width = 0.1) + 
    geom_point() + 
    facet_grid(.~exotic) +
    labs(x = "Estimate") +
    theme_bw() +
    theme(
      axis.title.y = element_blank() 
    )


ggsave(
  plot_glm_pant_lost_growth,
  file = "output/plot_glm_plant_lost_growth.png",
  width = 120, height = 60, units = "mm", dpi = 500)
```





Character of result of GLM of growth form

```{r}
cha_glm_plant_lost_growth <-
  bind_rows(
    tidy(glm_plant_lost$multi_grow_char[[1]]),
    tidy(glm_plant_lost$multi_grow_char[[2]])
    ) %>% 
  mutate(
    exotic = fct_inorder(rep(c("Exotic species","Native species"), each = 3)),
    growth = fct_inorder(rep(c("Annual","Perennial","Tree"), 2))
    )
```

Plot for result of glm of growth form

```{r}
box_plant_lost_growth <- 
  df_plant_lost %>% 
    group_by(species, growth, exotic) %>% 
    dplyr::summarise(lost = mean(lost), .groups = "drop") %>% 
    dplyr::filter(!is.na(growth)) %>% 
    dplyr:: mutate(
      exotic = recode_factor(
        exotic,
        "Native" = "Native species",
        "Exotic" = "Exotic species"),
      growth = recode(
        growth,
        "annual" = "Annual",
        "perennial" = "Perennial",
        "tree" = "Tree")
        ) %>% 
  
    # plot
    ggplot() +
    geom_boxplot(aes(x = growth, y = lost, fill = exotic)) + 
    facet_grid(~exotic) +
    geom_text(
      data = cha_glm_plant_lost_growth, 
      aes(x = growth, y = 1.05, label = letters),
      position = position_dodge(width = 1), size = 3) +
    labs(
      x = "Growth form",
      y = "Probability of species loss",
      fill = "Growth form"
      ) +
    theme_classic(base_size = 8) +
    scale_fill_simpsons() +
    #scale_fill_brewer(palette = "Paired") +
    theme(
      panel.grid = element_blank(),
      legend.position = "none",
      strip.background = element_blank(),
      strip.text = element_text(size = 8)
      )

box_plant_lost_growth
```

```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  box_plant_lost_growth,
  file = "output/box_plant_lost_growth.png",
  width = 100, height = 60, units = "mm", dpi = 500)
```

#### Dispersal type
```{r}
plot_glm_pant_lost_disper <- glm_plant_lost %>% 
  dplyr::select(exotic, multi_disper_conf) %>% 
  unnest(multi_disper_conf) %>% 
  dplyr::mutate(
    exotic = recode_factor(
      exotic,
      "Native" = "Native spcies",
      "Exotic" = "Exotic species"
      )
    ) %>% 
  
  ggplot(aes(x = estimate, y = contrast, xmin = conf.low, xmax = conf.high)) +
    geom_errorbar(width = 0.1) + 
    geom_point() + 
    facet_grid(.~exotic) +
    labs(x = "Estimate") +
    theme_bw() +
    theme(
      axis.title.y = element_blank() 
    )


ggsave(
  plot_glm_pant_lost_disper,
  file = "output/plot_glm_plant_lost_disper.png",
  width = 120, height = 60, units = "mm", dpi = 500)
```



Character of result of GLM of dispersal type

```{r}
cha_glm_plant_lost_disper <-
  bind_rows(
    tidy(glm_plant_lost$multi_disper_char[[1]]),
    tidy(glm_plant_lost$multi_disper_char[[2]])
    ) %>% 
  mutate(
    exotic = fct_inorder(rep(c("Exotic species","Native species"), each = 4)),
    growth = fct_inorder(rep(c("Animal","Gravity","Water","Wind"), 2))
    )
```

Plot for result of glm of growth form

```{r}
box_plant_lost_disper <- 
  df_plant_lost %>% 
    group_by(species, disper, exotic) %>% 
    dplyr::summarise(lost = mean(lost), .groups = "drop") %>% 
    dplyr::filter(!is.na(disper)) %>% 
    dplyr:: mutate(
      exotic = recode_factor(
        exotic,
        "Native" = "Native species",
        "Exotic" = "Exotic species"),
      disper = recode(
        disper,
        "wind" = "Wind",
        "animal" = "Animal",
        "water" = "Water",
        "gravity" = "Gravity")
        ) %>% 
  
    # plot
    ggplot() +
    geom_boxplot(aes(x = disper, y = lost, fill = exotic)) + 
    facet_grid(~exotic) +
    geom_text(
      data = cha_glm_plant_lost_disper, 
      aes(x = growth, y = 1.05, label = letters),
      position = position_dodge(width = 1), size = 3) +
    labs(
      x = "Dispersal type",
      y = "Probability of species loss"
      ) +
    theme_classic(base_size = 8) +
    scale_fill_simpsons() +
    theme(
      panel.grid = element_blank(),
      legend.position = "none",
      strip.background = element_blank(),
      strip.text = element_text(size = 8)
      )

box_plant_lost_disper
```

```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  box_plant_lost_disper,
  file = "output/box_plant_lost_disper.png",
  width = 100, height = 60, units = "mm", dpi = 500)
```

#### Environment
Plot for the result of GLM of environmental variables
```{r}
theme_set(theme_bw(base_size = 9))

# Year
p_glm_plant_lost_year <- 
  ggplot() +
    geom_point(
      data = df_plant_lost ,
      aes(x = year, y = lost, color = fct_rev(exotic))) +
    geom_smooth(
      data = glm_plant_lost %>% 
        unnest(pred_year),
      aes(x = x, y = predicted, color = fct_rev(exotic)),
      method = "glm", se = FALSE,
      method.args = list(family = "binomial")) +
    labs(
      x = "Years between surveys",
      y = "Probability of species loss"
      ) +
      scale_color_simpsons() +
      theme(
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 9),
        legend.position = c(0.25, 0.75),
        legend.background = element_blank(),
        legend.title = element_blank()
        )
  

# Area
p_glm_plant_lost_area <- 
  ggplot() +
    geom_point(
      data = df_plant_lost,
      aes(x = area/10000, y = lost, color = fct_rev(exotic))) +
    geom_smooth(
      data = glm_plant_lost %>% 
        dplyr::filter(exotic == "Native") %>% 
        unnest(pred_area),
      aes(x/10000, predicted, color = fct_rev(exotic)),
      method = "glm", se = FALSE,
      method.args = list(family = "binomial")) +
    labs(
      x = "Area (ha)",
      y = ""
      ) +
    scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )


# Green rate
p_glm_plant_lost_green <- 
  ggplot() +
    geom_point(
      data = df_plant_lost,
      aes(x = green_rate, y = lost, color = fct_rev(exotic))) +
    geom_smooth(
      data = glm_plant_lost %>% 
        unnest(pred_green),
      aes(x, predicted, color = fct_rev(exotic)),
      method = "glm", se = FALSE,
      method.args = list(family = "binomial")) +
      labs(
        x = "Green rate (%)",
        y = ""
      ) +
      scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )



# Combine
p_glm_plant_lost_year + p_glm_plant_lost_area + p_glm_plant_lost_green
```
```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  p_glm_plant_lost_year + p_glm_plant_lost_area + p_glm_plant_lost_green,
  file = "output/plot_glm_plant_lost_env.png",
  width = 170, height = 60, units = "mm", dpi = 500)
```


## Colonized species

Estimate a number of colonized species

```{r}
df_plant_colo <- df_plant %>%
  pivot_wider(names_from = time, values_from = value) %>% 
  dplyr::mutate(colo = if_else(now == 1 & past == 0, 1, 0)) %>% 
  dplyr::left_join(env, by = "site") %>% 
  dplyr::filter(!is.na(exotic) & now == 1) %>% 
  dplyr::mutate(across(where(is.character), as.factor))
```

Confirmation of number of colonized species across dispersal type
```{r}
df_plant_colo%>% 
  dplyr::group_by(exotic, disper) %>% 
  dplyr::summarise(n = sum(colo)/n()) %>% 
  na.omit()
```

### GLM

Relationships between colonized species and trait, environmental variables

```{r}
glm_plant_colo <- df_plant_colo %>% 
    dplyr::mutate(across(where(is.character), as.factor)) %>% 
    group_nest(exotic) %>% 
    dplyr::mutate(
      
      # GLM
      fit = map(
        data, 
        ~glm(colo ~ year + area + green_rate + growth + disper, 
             family = "binomial", data = .)),
      
      # summary of GLM
      summary_fit = map(fit, ~tidy(.)),
      
      
      # Growth form
      multi_grow = map(fit, ~glht(., linfct = mcp(growth = "Tukey"))),
      multi_grow_summary = map(multi_grow, ~tidy(.)),
      multi_grow_conf = map(multi_grow, ~confint(.) %>% tidy()),
      multi_grow_char = map(multi_grow, ~cld(., decreasing = F)),
      
      
      # Dispersal type
      multi_disper = map(fit, ~glht(., linfct = mcp(disper = "Tukey"))),
      multi_disper_summary = map(multi_disper, ~tidy(.)),
      multi_disper_conf = map(multi_disper, ~confint(.)%>% tidy()),
      multi_disper_char = map(multi_disper, ~cld(., decreasing = F)),
      
      
      # Environment
      pred_year = map(fit, ~ggpredict(., terms = "year")),
      pred_area = map(fit, ~ggpredict(., terms = "area")),
      pred_green = map(fit, ~ggpredict(., terms = "green_rate"))
      ) 
```



Table for result of GLM

```{r}
tb_glm_plant_colo <- 
  glm_plant_colo %>% 
    dplyr::select(exotic, summary_fit) %>% 
    unnest(summary_fit) %>% 
    dplyr::filter(term != "(Intercept)") %>% 
    dplyr::mutate(
      across(where(is.numeric),
             ~ifelse(. > -0.001 & . < 0.001,
                      formatC(., digits = 3, format = "e"),
                      formatC(., digits = 3, format = "fg"))
      )) %>% 
    dplyr::rename(
      Nativity = exotic,
      "Explanatory variables" = term,
      Estimate = estimate,
      Std.error = std.error,
       Statistic = statistic) %>% 
    print()
```
```{r eval=FALSE, include=FALSE}
write.csv(tb_glm_plant_colo, "output/glm_colo_plant.csv")
```


Partial r squared
```{r }

# All species
glm_plant_colo <- 
  glm(colo ~ year + area + green_rate + growth + disper, 
      family = "binomial", data = df_plant_colo)

rsq_glm_plant_lost <- rsq.partial(glm_plant_colo, adj = TRUE)



# Native species
df_plant_colo_native <-
  df_plant_colo %>% 
    dplyr::filter(exotic == "Native" & !is.na(growth) & !is.na(disper)) 

glm_plant_colo_native <- 
  glm(colo ~ year + area + green_rate + growth + disper, 
      family = "binomial", data = df_plant_colo_native)

rsq_glm_plant_colo_native <- rsq.partial(glm_plant_colo_native, adj = TRUE)




# Exotic species
df_plant_colo_exotic <-
  df_plant_colo %>% 
    dplyr::filter(exotic == "Exotic" & !is.na(growth) & !is.na(disper)) 

glm_plant_colo_exotic <- 
  glm(colo ~ year + area + green_rate + growth + disper, 
      family = "binomial", data = df_plant_colo_exotic)

rsq_glm_plant_colo_exotic <- rsq.partial(glm_plant_colo_exotic, adj = TRUE)


df_rsq_glm_plant_colo <- 
  bind_rows(
    rsq_glm_plant_lost[2:3],
    rsq_glm_plant_colo_native[2:3],
    rsq_glm_plant_colo_exotic[2:3]
    ) %>% 
    dplyr::mutate(Nativity = rep(c("All", "Native","Exotic"), each = 5)) %>% 
  print()
```
```{r eval=FALSE, include=FALSE}
write.csv(df_rsq_glm_plant_colo, "output/rsq_glm_plant_colo.csv")
```

### Plot
#### Growth form

```{r}
plot_glm_pant_colo_growth <- glm_plant_colo %>% 
  dplyr::select(exotic, multi_grow_conf) %>% 
  unnest(multi_grow_conf) %>% 
  dplyr::mutate(
    exotic = recode_factor(
      exotic,
      "Native" = "Native spcies",
      "Exotic" = "Exotic species"
      )
    ) %>% 
  
  ggplot(aes(x = estimate, y = contrast, xmin = conf.low, xmax = conf.high)) +
    geom_errorbar(width = 0.1) + 
    geom_point() + 
    facet_grid(.~exotic) +
    labs(x = "Estimate") +
    theme_bw() +
    theme(
      axis.title.y = element_blank() 
    )


ggsave(
  plot_glm_pant_lost_growth,
  file = "output/plot_glm_plant_colo_growth.png",
  width = 120, height = 60, units = "mm", dpi = 500)
```





Character of result of GLM of growth form

```{r}
cha_glm_plant_colo_growth <-
  bind_rows(
    tidy(glm_plant_colo$multi_grow_char[[1]]),
    tidy(glm_plant_colo$multi_grow_char[[2]])
    ) %>% 
  mutate(
    exotic = fct_inorder(rep(c("Native species","Exotic species"), each = 3)),
    growth = fct_inorder(rep(c("Annual","Perennial","Tree"), 2))
    )
```







Plot for result of glm of growth form

```{r}
box_plant_colo_growth <- 
  df_plant_colo %>% 
    group_by(species, growth, exotic) %>% 
    dplyr::summarise(colo = mean(colo), .groups = "drop") %>% 
    dplyr::filter(!is.na(growth)) %>% 
    dplyr:: mutate(
      exotic = recode_factor(
        exotic,
        "Native" = "Native species",
        "Exotic" = "Exotic species"),
      growth = recode(
        growth,
        "annual" = "Annual",
        "perennial" = "Perennial",
        "tree" = "Tree")
        ) %>% 
  
    # plot
    ggplot() +
    geom_boxplot(aes(x = growth, y = colo, fill = exotic)) + 
    facet_grid(~exotic) +
    geom_text(
      data = cha_glm_plant_colo_growth, 
      aes(x = growth, y = 1.05, label = letters),
      position = position_dodge(width = 1), size = 3) +
    labs(
      x = "Growth form",
      y = "Probability of species colonization",
      fill = "Growth form"
      ) +
    theme_classic(base_size = 8) +
    scale_fill_simpsons() +
    #scale_fill_brewer(palette = "Paired") +
    theme(
      panel.grid = element_blank(),
      legend.position = "none",
      strip.background = element_blank(),
      strip.text = element_text(size = 8)
      )

box_plant_colo_growth
```

```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  box_plant_colo_growth,
  file = "output/box_plant_colo_growth.png",
  width = 100, height = 60, units = "mm", dpi = 500)
```


#### Dispersal type

```{r}
plot_glm_pant_colo_disper <- glm_plant_colo %>% 
  dplyr::select(exotic, multi_disper_conf) %>% 
  unnest(multi_disper_conf) %>% 
  dplyr::mutate(
    exotic = recode_factor(
      exotic,
      "Native" = "Native spcies",
      "Exotic" = "Exotic species"
      )
    ) %>% 
  
  ggplot(aes(x = estimate, y = contrast, xmin = conf.low, xmax = conf.high)) +
    geom_errorbar(width = 0.1) + 
    geom_point() + 
    facet_grid(.~exotic) +
    labs(x = "Estimate") +
    theme_bw() +
    theme(
      axis.title.y = element_blank() 
    )


ggsave(
  plot_glm_pant_colo_disper,
  file = "output/plot_glm_plant_colo_disper.png",
  width = 120, height = 60, units = "mm", dpi = 500)
```



Character of result of GLM of dispersal type

```{r}
cha_glm_plant_colo_disper <-
  bind_rows(
    tidy(glm_plant_colo$multi_disper_char[[1]]),
    tidy(glm_plant_colo$multi_disper_char[[2]])
    ) %>% 
  dplyr::mutate(
    exotic = fct_inorder(rep(c("Exotic species","Native species"), each = 4)),
    growth = fct_inorder(rep(c("Animal","Gravity","Water","Wind"), 2))
    )
```

Plot for result of glm of growth form

```{r}
box_plant_colo_disper <- 
  df_plant_colo %>% 
    group_by(species, disper, exotic) %>% 
    dplyr::summarise(colo = mean(colo), .groups = "drop") %>% 
    dplyr::filter(!is.na(disper)) %>% 
    dplyr:: mutate(
      exotic = recode_factor(
        exotic,
        "Native" = "Native species",
        "Exotic" = "Exotic species"),
      disper = recode(
        disper,
        "wind" = "Wind",
        "animal" = "Animal",
        "water" = "Water",
        "gravity" = "Gravity")
        ) %>% 
  
    # plot
    ggplot() +
    geom_boxplot(aes(x = disper, y = colo, fill = exotic)) + 
    facet_grid(~exotic) +
    geom_text(
      data = cha_glm_plant_colo_disper, 
      aes(x = growth, y = 1.05, label = letters),
      position = position_dodge(width = 1), size = 3) +
    labs(
      x = "Dispersal type",
      y = "Probability of species colonization"
      ) +
    theme_classic(base_size = 8) +
    scale_fill_simpsons() +
    theme(
      panel.grid = element_blank(),
      legend.position = "none",
      strip.background = element_blank(),
      strip.text = element_text(size = 8)
      )

box_plant_colo_disper
```

```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  box_plant_colo_disper,
  file = "output/box_plant_colo_disper.png",
  width = 100, height = 60, units = "mm", dpi = 500)
```


#### Environment
Plot for the result of GLM of environmental variables
```{r}
theme_set(theme_bw(base_size = 9))

# Year
p_glm_plant_colo_year <- 
  ggplot() +
    geom_point(
      data = df_plant_colo,
      aes(x = year, y = colo, color = fct_rev(exotic))) +
    geom_smooth(
      data = glm_plant_colo %>% 
        unnest(pred_year),
      aes(x = x, y = predicted, color = fct_rev(exotic)),
      method = "glm", se = FALSE,
      method.args = list(family = "binomial")) +
    labs(
      x = "Years between surveys",
      y = "Probability of species colonization"
      ) +
      scale_color_simpsons() +
      theme(
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 9),
        legend.position = c(0.25, 0.75),
        legend.background = element_blank(),
        legend.title = element_blank()
        )
  

# Area
p_glm_plant_colo_area <- 
  ggplot() +
    geom_point(
      data = df_plant_colo,
      aes(x = area/10000, y = colo, color = fct_rev(exotic))) +
    geom_smooth(
      data = glm_plant_lost %>% 
        dplyr::filter(exotic == "Native") %>%  
        unnest(pred_area),
      aes(x/10000, predicted, color = fct_rev(exotic)),
      method = "glm", se = FALSE,
      method.args = list(family = "binomial")) +
    labs(
      x = "Area (ha)",
      y = ""
      ) +
    scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )


# Green rate
p_glm_plant_colo_green <- 
  ggplot() +
    geom_point(
      data = df_plant_colo,
      aes(x = green_rate, y = colo, color = fct_rev(exotic))) +
    geom_smooth(
      data = glm_plant_colo %>% 
        dplyr::filter(exotic == "Native") %>% 
        unnest(pred_green),
      aes(x, predicted, color = fct_rev(exotic)),
      method = "glm", se = FALSE,
      method.args = list(family = "binomial")) +
      labs(
        x = "Green rate (%)",
        y = ""
      ) +
      scale_color_simpsons() +
      theme(
        axis.title.y = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none"
      )



# Combine
p_glm_plant_colo_year + p_glm_plant_colo_area + p_glm_plant_colo_green
```
```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  p_glm_plant_colo_year + p_glm_plant_colo_area + p_glm_plant_colo_green,
  file = "output/plot_glm_plant_colo_env.png",
  width = 170, height = 60, units = "mm", dpi = 500)
```



# Species composition

```{r}
df_comp_plant <- 
  df_plant_trait %>% 
  dplyr::select(site:value, exotic) %>% 
  dplyr::filter(!is.na(exotic)) %>% 
  pivot_wider(names_from = species, values_from = value, values_fill = 0)  


# data-frame for all, native, and exotic species
comp_plant <- df_comp_plant %>% 
  dplyr::group_by(site, time) %>% 
  dplyr::summarise(across(where(is.numeric), sum), .groups = "drop") %>% 
  dplyr::mutate(exotic = "All species") %>% 
  dplyr::select(site, time, exotic, everything()) %>% 
  dplyr::bind_rows(df_comp_plant) %>% 
  nest_by(exotic) %>% 
  ungroup() %>% 
  dplyr::mutate(
    comp = map(data, 
               ~dplyr::select(., where(~ is.numeric(.) && sum(.) != 0))),
    
    # NMDS
    nmds = map(comp, 
               ~metaMDS(., dist = "jac", k = 2, trace = F, trymax = 999)),
    
    # NMDS axis score
    nmds_score = map(nmds, ~scores(.) %>% pluck(1) %>% as.data.frame()),
    
    # NMDS stress value
    nmds_stress = map_dbl(nmds, ~.$stress),
    
    # permanova
    permanova = map(data, ~adonis2(.[,-1:-2] ~ .$time, method = "jac")),
    
    
    # p-value from permanova
    pval = map(permanova, ~as.tibble(.) %>% dplyr::select("Pr(>F)")),
    
    
    exotic = recode_factor(exotic, 
                           "Native" = "Native species",
                           "Exotic" = "Exotic species")
    ) 
```

## Permanova

```{r}
df_label_plant <-
  comp_plant %>% 
    dplyr::select(exotic, nmds_stress, pval) %>% 
    unnest(pval) %>% 
    na.omit() %>% 
    dplyr::mutate(
      across(where(is.numeric), ~round(., digits = 3)),
      time = "Past",
      stressvalue = "Stress value:",
      permanova_cha = "Permanova: p = "
      ) %>% 
    unite(stress, stressvalue, nmds_stress, remove = F, sep = "") %>% 
    unite(permanova, permanova_cha, "Pr(>F)", remove = F, sep = "") %>% 
    print()
```

Plot for result of nmds

```{r}
p_nmds_plant <- 
  comp_plant %>% 
  dplyr::select(exotic, nmds_score) %>% 
  unnest(nmds_score) %>% 
  dplyr::mutate(
    time = rep(c("Past", "Present"), 45),
    time = factor(time, levels = c("Past", "Present")),
    exotic = factor(
      exotic, 
      levels = c("All species", "Native species", "Exotic species"))
    ) %>% 
  
    ggplot(aes(x = NMDS1, y = NMDS2, color = time)) +
      geom_point() +
      stat_ellipse(
        aes(group = time, fill = time), 
        alpha = 0.1,
        geom = "polygon") +
  
      facet_wrap(.~exotic) +
      
  
      # Stress value
      geom_text(
        data = df_label_plant, aes(label = stress), 
        x = 0.85, y = 1.2, size = 2, color = "grey50", hjust = 0) +
      
      # P value from Permanova
      geom_text(
        data = df_label_plant, aes(label = permanova), 
        x = 0.85, y = 1.1, size = 2, color = "grey50", hjust = 0) +
  
      theme_bw(base_size = 9) +
      theme(
        panel.grid = element_blank(),
        legend.position = c(0.08, 0.9),
        legend.title = element_blank(),
        legend.background = element_blank(),
        strip.background = element_blank()
        )

p_nmds_plant
```

```{r eval=FALSE, include=FALSE}
# For save
ggsave(
  p_nmds_plant,
  file = "output/plot_nmds_plant.png",
  width = 170, height = 90, units = "mm", dpi = 500)
```
